---
title: "code.qmd"
author: "Kyra Baranick"
format: html
editor: visual
---

[My GitHub Repository](baranickkyra/ENVS-193DS_homework-03:%20Homework%20Three%20for%20ENVS193DS)

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(janitor)
library(readxl)

salinity <- read.csv(here("data","salinity-pickleweed.csv"))
Kyra_data <- read.csv(here("data","personal_data.csv"))

```

# Problem 1

### a.

I could do a linear regression test or a Pearson's correlation test. Pearson’s correlation measures the stregeth and direction of a linear association between the two variables, while linear regression quantifies the relationship (1 predictor and 1 response) by estimating how biomass changes with salinity (uses slope and intercept).

### b.

```{r}
#| message: false
#| warning: false
ggplot(salinity, aes(x = salinity_mS_cm, y = pickleweed)) +
  geom_point(color = "darkorange") +
  geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
  labs(
    x = "Soil salinity (mS/cm)",
    y = "Pickleweed biomass (g)",
    title = "Relationship between soil salinity and pickleweed biomass"
  ) +
  theme_minimal()

```

### c.

# Problem 2

### a.

```{r}
ggplot(Kyra_data, # making a ggplot using the data from Proj1
       aes(x = Weekend.or.Weekday,   # using weekend or weekday as the categorical xaxis
           y = Pickups, # using pickups data for the y axis
           fill = Weekend.or.Weekday)) + # filling with different colors for weekendand weekday
  geom_boxplot(alpha = 0.2) + # adjusting the transparency
  geom_jitter(width = 0.2, # giving the points a slight horizontal jitter
              height = 0, # NO vertical jitter
              alpha = 0.8, # making points semi-transparent for readability
              shape = 21) + # picking a different shape
  labs(
    title = "Phone Pickup Count: Weekends vs Weekdays", # creating the title ofthe graph
    subtitle = "Most recent observation: February 18, 2026",
    x = "Day Type", # labeling the x axis
    y = "Pickup Count" # labeling the y axis
  ) +
  scale_fill_manual(values = c( # filling with colors otherthen the default
    "weekday" = "red",
    "weekend" = "purple"
  )) +
  
  theme_classic() + # changed the theme form the default
  guides(fill = "none") # removing the legend

```

```{r}
ggplot(Kyra_data, # creating a plot using ggplot 
       aes(x = `min.of.sleep.the.night.before`, # using min of sleep as the x axis 
           y = Pickups)) + # using pickups as the y-axis
  
  geom_point(color = "#1B9E77",   # nondefault green 
             size = 3 # adjusting the size for visibilty 
             ) + #
  
  geom_smooth(method = "lm",
              se = TRUE,           # shows confidence band
              color = "#D95F02",   # non-default orange
              linewidth = 1.2) +   # widening the line just to do soem extra messing around with the visual
  
  labs(
    title = "Phone Pickups vs. Minutes of Sleep", # adding a title
    subtitle = "Most recent observation: February 18, 2026", # putting my most recent obs as the subtitle 
    x = "Minutes of sleep the night before (min)", # labeling my x-axis
    y = "Phone pickup count (number of pickups)"    # labeling my y-axis
  ) +

  theme_classic() + # changing the theme form the gghplot default 

  theme(
    panel.grid = element_blank(),        # removes all gridlines
    panel.background = element_blank(),  # removes grey panel background
    plot.background = element_blank(),   # removes outer plot background
    axis.line = element_line(color = "black"),  # keeps clean black axis lines
    axis.ticks = element_line(color = "black")  # keeps ticks visible and clean
  )

```

### b. 

FIGURE 1: The x-axis (Day Type) shows whether the day is a weekday or weekend, and the y-axis (Pickup Count) shows the number of phone pickups. The red box represents weekdays and the purple box represents weekends. Dots show individual daily observations, pink being weekdays and purple for weekends. The black line inside each box is the median, the box edges show the interquartile range (middle 50% of values), and the vertical lines (whiskers) show the range of non-outlier values. The subtitle indicates the most recent observation date (February 18, 2026).

FIGURE 2 : The x-axis (Minutes of sleep the night before) shows sleep duration in minutes, and the y-axis (Phone pickup count) shows the number of daily phone pickups. Green dots represent individual daily observations. The red line shows the fitted linear trend between sleep and phone pickups, and the gray shaded band represents the confidence interval around the trend line. The subtitle indicates the most recent observation date (February 18, 2026).

### c. 
